---
title: 'Homework #2'
author: "Jaxon Stuhr"
date: "2022-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(corrplot)
library(ggthemes)
library(grid)
library(here)
```


Load in the abalone dataset
```{r}
abalone_raw = read_csv(here("data", "abalone.csv"))
set.seed(23)
```

# Question 1
Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no age variable in the data set. Add age to the data set.

Assess and describe the distribution of age.

```{r}
abalone = abalone_raw %>% 
# add age variable to data
  mutate(age = rings + 1.5) %>% 
  select(-rings)

par(mfrow=c(1,2))

ggplot(abalone, aes(x = age)) +
  geom_histogram()

ggplot(abalone, aes(sample = age)) +
  stat_qq() +
  stat_qq_line(lwd = 1)

mean(abalone$age)
median(abalone$age)
```

The distribution of age appears relatively normal around a median of 10.5 years, however, there is an exceptionally large tail of very old abalone, which brings the mean age up to 11.4 years, and appears highly non-normal in its distribution. 

# Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

```{r}
abalone_split <- initial_split(abalone, prop = 0.75,
                                strata = age)

abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)
```

# Question 3

Using the training data, create a recipe predicting the outcome variable, age, with all other predictor variables. Note that you should not include rings to predict age. Explain why you shouldn’t use rings to predict age.

We should not use rings to predict age because our definition of age is directly based on the number of rings. We want to understand weather the age of abalone can be predicted from more readily available indicators.

```{r}
# generate recipe age as a function of all predictors
abalone_recipe <- recipe(age ~ ., data = abalone_train) %>% 
  # dummy vars for nominal predictors
  step_dummy(type) %>% 
  # set up interaction terms
  step_interact(terms = ~ starts_with("type"):shucked_weight) %>% 
  step_interact(terms = ~ longest_shell:diameter) %>% 
  step_interact(terms = ~ shucked_weight:shell_weight) %>%
  # center
  step_center(all_numeric_predictors()) %>%
  # scale
  step_scale(all_numeric_predictors())
```


# Question 4

Create and store a linear regression object using the "lm" engine.

```{r}
lm_model <- linear_reg() %>% 
  set_engine("lm")
```

# Question 5

Now:

set up an empty workflow,
add the model you created in Question 4, and
add the recipe that you created in Question 3.

```{r}
# set up empty workflow
lm_wflow <- workflow() %>% 
  # add model
  add_model(lm_model) %>% 
  # add recipe
  add_recipe(abalone_recipe)
```

# Question 6

Use your fit() object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}
lm_fit <- fit(lm_wflow, abalone_train)

sample_abalone = abalone[0,] %>% 
  select(-age) %>% 
  add_row(type = "F",  longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1)

sample_prediction <- predict(lm_fit, new_data = sample_abalone)
```

Predicted Age: 24.8 years





